### 2013-03-28 doc/sosp13/tasks.tex
tasks stuff

The basic unit of execution in \Grappa is a {\em task}. Each hardware core has a single operating system thread pinned to it. When tasks are ready to execute, they are mapped to a {\em worker}, which is akin to a user-level thread. \Grappa keeps a large pool of workers per hardware core. Below we discuss the implementation of our task management support and then describe how applications expose parallelism to the \Grappa runtime.

@@

@@

\noindent {\it private argument}, often a loop index;

\noindent {\it shared argument}, typically data shared amongst a group of
tasks, or the number of loop iterations to be performed by this task;

\noindent {\it synchronization argument}, often used to determine when all
tasks that are part of a loop have finished, in the form of a global pointer
to a synchronization object allocated at the core that spawned a group of
tasks.

@@

\paragraph{Workers} Workers execute application tasks.  A worker is simply a collection of status bits and a stack, allocated at a particular core.  When a task is ready to be executed it is assigned to a worker.  Once a task is mapped to a worker it stays with that worker until it completes.

%% Note sure this is even true.  And it's confusing and not relevant regardless. -M
% There are two types of workers: (1) user worker; and (2) system worker. 
%System workers execute runtime management activities such as communication and statistics collection. User workers and system workers might have different scheduling priorities. System workers execute periodically and in idle times, to guarantee timely communication services.

\paragraph{Scheduling} 
During execution, a worker yields control of its core whenever it performs a
long-latency operation, allowing the processor to remain busy while waiting
for the operation to complete. In addition, a programmer can direct scheduling
explicitly via the \Grappa API calls shown in Figure~\ref{fig:scheduling}. To
minimize yield overhead, the \Grappa scheduler operates entirely in user-space
and does little more than store register state of one worker and load that of
another.

@@

Each core in a \Grappa system has its own independent scheduler.  The scheduler has a single queue of active workers:

\noindent {\it ready worker queue}, a FIFO queue of workers that are
  matched with workers and are ready to execute;

Each scheduler also has three queues of tasks waiting to be assigned a worker:

\noindent {\it deadline task queue}, a priority queue of tasks that are executed according to task-specific deadline constraints.

\noindent {\it private task queue}, a FIFO queue of tasks that must run on this core;

\noindent {\it public task queue},  a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool.

@@

\paragraph{Context switching} 
\Grappa context switches between workers non-preemptively. As with other
cooperative multithreading systems, we treat context switches as function
calls, saving and restoring only the callee-saved state as specified in the
x86-64 ABI\~cite{amd64:abi:2012}. This involves saving six general-purpose
64-bit registers and the stack pointer, as well as the 16-bit x87 floating
point control word and the SSE context/status register. Thus, the minimum
amount of state a cooperative context switch routine must save, according to
the ABI, is 62~bytes.

Since \Grappa keeps a very large number of active workers, their context data
will not fit in cache. Therefore, the scheduler has to perform very careful
prefetching in order to reduce context switch cost to a minimum. We determined
empirically that prefetching the fourth worker in the scheduling order is
sufficient. We also prefetch three cache lines from the stack for each worker.
This makes context switching effectively free of cache misses even to hundreds
of thousands of threads. We provide a detailed analysis of our context switch
performance in Section~\ref{eval:ctx}.

>>>>

Below we discuss the implementation of our task management support and then
describe how applications expose parallelism to the \Grappa runtime.

@@

The basic unit of execution in \Grappa is a {\em task}. Each hardware core has
a single operating system thread pinned to it. When tasks are ready to
execute, they are mapped to a {\em worker}, which is akin to a user-level
thread. % \Grappa keeps a large pool of workers per hardware core.

@@

\begin{itemize}
\item {\it private argument}, often a loop index;

\item {\it shared argument}, typically data shared amongst a group of
tasks, or the number of loop iterations to be performed by this task;

\item {\it synchronization argument}, often used to determine when all
tasks that are part of a loop have finished, in the form of a global pointer
to a synchronization object allocated at the core that spawned a group of
tasks.
\end{itemize}

@@

\paragraph{Workers} Workers execute application and system (e.g.,
communication) tasks. A worker is simply a collection of status bits and a
stack, allocated at a particular core. When a task is ready to be executed it
is assigned to a worker. Once a task is mapped to a worker it stays with that
worker until it completes.

\paragraph{Scheduling} During execution, a worker yields control of its core
whenever it performs a long-latency operation, allowing the processor to
remain busy while waiting for the operation to complete. In addition, a
programmer can direct scheduling explicitly via the \Grappa API calls shown in
Figure~\ref{fig:scheduling}. To minimize yield overhead, the \Grappa scheduler
operates entirely in user-space and does little more than store state of one
worker and load that of another.

@@

Each core in a \Grappa system has its own independent scheduler. The scheduler
has a single FIFO queue of active \emph{workers} ready to execute, the {\it
ready worker queue}. Each scheduler also has three queues of \emph{tasks}
waiting to be assigned a worker:

\begin{itemize}

\item {\it deadline task queue}, a priority queue of tasks that are executed according to task-specific deadline constraints;

\item {\it private task queue}, a FIFO queue of tasks that must run on
this core and therefore is not subject to stealing;

\item {\it public task queue},  a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool.

\end{itemize}

@@

\paragraph{Context switching} 
\Grappa context switches between workers non-preemptively. As with other
cooperative multithreading systems, we treat context switches as function
calls, saving and restoring only the callee-saved state as specified in the
x86-64 ABI~\cite{amd64:abi:2012}. This involves saving six general-purpose
64-bit registers and the stack pointer, as well as the 16-bit x87 floating
point control word and the SSE context/status register. Thus, the minimum
amount of state a cooperative context switch routine must save, according to
the ABI, is 62~bytes.

Since \Grappa keeps a very large number of active workers, their context data
will not fit in cache. Therefore, the scheduler has to perform very careful
prefetching in order to reduce context switch cost to a minimum. We determined
empirically that prefetching the fourth worker in the scheduling order is
sufficient. We also prefetch three cache lines from the stack for each worker.
This makes context switching effectively free of cache misses even to hundreds
of thousands of threads. We provide an analysis of our context switch
performance in Section~\ref{eval:basic}.
